### 2016-12-24 ack.tex

I take this opportunity to express my profound gratitude to the people who have helped and supported me throughout my PhD. I thank my guide, Prof. Shalabh Bhatnagar, for his constant and valuable suggestions during my research. Without his constant guidance and support, this thesis would not have been possible. Next I thank Prof. Borkar for several interesting and useful discussions on our collaborative work. I thank Sunil, Hemanth and Sindhu for enthusiastically collaborating on the research problems. I thank Prof. Narahari, Electrical Sciences Divisional Chairman, who has always been keen to know how my thesis work is progressing. I am grateful to all the professors of CSA Dept for sharing their knowledge and for their encouragement.  I, also, thank the CSA Dept office staff for their invaluable support.  I would like to particularly acknowledge Mrs. Suguna and Mrs. Lalitha who promptly helped me on many occasions. Last but not the least, I express my heartfelt gratitude to the Almighty, my parents, younger sister, relatives and friends for their love and blessings.