### 2017-01-05 ack.tex
experiments added

I take this opportunity to express my profound gratitude to the people who have helped and supported me throughout my M.Sc(engg). I thank my guide, Prof. Shalabh Bhatnagar, for his constant and valuable suggestions during my research. Without his constant guidance and support, this thesis would not have been possible. Next I thank Prashanth L.A. for several interesting and useful discussions on our collaborative work. I thank Chandramouli K. for his help during the course work. I am grateful to all the professors of CSA Dept for sharing their knowledge and for their encouragement.  I, also, thank the CSA Dept office staff for their invaluable support. Last but not the least, I express my heartfelt gratitude to  my parents, sister, brother-in-law, relatives and friends for their love and blessings.