### 2017-01-05 secorder-rdsa-ih.tex

\chapter{Second-order RDSA with improved hessian estimation (2RDSA-IH)}
\label{sec:2rdsa-ih}
The second-order RDSA with improved Hessian estimate performs an update iteration as follows:
\begin{align}
\label{eq:e2rdsa}
\theta_{n+1} = & \; \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n), \\
\overline H_n = & \; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),\label{eq:2rdsa-H}
\end{align}
where $\widehat\nabla f(\theta_n)$ is the estimate of $\nabla f(\theta_n)$, 
 %and this corresponds to  the uniform  and   asymmetric Bernoulli variant gradient estimates proposed in \cite{prashanth2015rdsa}.
$\overline H_n$ is an estimate of the true Hessian ${\nabla}^2 f(\cdot)$, $\Upsilon(\cdot)$ projects any matrix onto the set of positive definite matrices and $\{a_n, n\ge 0\}$ is a step-size sequence that satisfies standard stochastic approximation conditions. There are standard procedures such as Cholesky factorization, see \cite{bert22}, for projecting a given square matrix to set of positive definite matrices. Moreover, in the vicinity of a local minimum, one expects the Hessian to be positive definite. In such a case, $\Upsilon$ will represent the identity operator.

The recursion \eqref{eq:e2rdsa} is identical to that in 2RDSA, while the Hessian estimation recursion \eqref{eq:2rdsa-H} differs as follows:\\
\begin{inparaenum}[\bfseries (i)]
\item  $\widehat \Psi_n$ is a zero-mean feedback term that reduces the error in Hessian estimate; and\\
\item $b_n$ is a general step-size that we optimize to improve the Hessian estimate.
\end{inparaenum}

On the other hand, $\widehat H_n$ is identical to that in 2RDSA, i.e., it estimates the true Hessian in each iteration using $3$ function evaluations. %Further, $\widehat H_0 = I$. 
For the sake of completeness, we provide below the construction for $\widehat\nabla f(\theta_n)$ and $\widehat H_n$ using asymmetric Bernoulli perturbations, before we present the feedback term that reduces the error in $\widehat H_n$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\algblock{PEval}{EndPEval}
%\algnewcommand\algorithmicPEval{\textbf{\em Function evaluation 1}}
% \algnewcommand\algorithmicendPEval{}
%\algrenewtext{PEval}[1]{\algorithmicPEval\ #1}
%\algrenewtext{EndPEval}{\algorithmicendPEval}
%
%\algblock{PEvalPrime}{EndPEvalPrime}
%\algnewcommand\algorithmicPEvalPrime{\textbf{\em Function evaluation 2}}
% \algnewcommand\algorithmicendPEvalPrime{}
%\algrenewtext{PEvalPrime}[1]{\algorithmicPEvalPrime\ #1}
%\algrenewtext{EndPEvalPrime}{\algorithmicendPEvalPrime}
%
%\algblock{PImp}{EndPImp}
%\algnewcommand\algorithmicPImp{\textbf{\em Gradient descent}}
% \algnewcommand\algorithmicendPImp{}
%\algrenewtext{PImp}[1]{\algorithmicPImp\ #1}
%\algrenewtext{EndPImp}{\algorithmicendPImp}
%
%\algtext*{EndPEval}
%\algtext*{EndPEvalPrime}
%\algtext*{EndPImp}
%%%%%%%%%%%%%%%%% alg-custom-block %%%%%%%%%%%%
%\algblock{PEvalPrimeDouble}{EndPEvalPrimeDouble}
%\algnewcommand\algorithmicPEvalPrimeDouble{\textbf{\em Function evaluation 3}}
% \algnewcommand\algorithmicendPEvalPrimeDouble{}
%\algrenewtext{PEvalPrimeDouble}[1]{\algorithmicPEvalPrimeDouble\ #1}
%\algrenewtext{EndPEvalPrimeDouble}{\algorithmicendPEvalPrimeDouble}
%\algtext*{EndPEvalPrimeDouble}
%
%\algblock{PImpNewton}{EndPImpNewton}
%\algnewcommand\algorithmicPImpNewton{\textbf{\em Newton step}}
% \algnewcommand\algorithmicendPImpNewton{}
%\algrenewtext{PImpNewton}[1]{\algorithmicPImpNewton\ #1}
%\algrenewtext{EndPImpNewton}{\algorithmicendPImpNewton}
%
%\algtext*{EndPImpNewton}
%
%%%%%%%%%%%%%%%%%%%%
%
%\begin{algorithm}[t]
%\begin{algorithmic}
%\State {\bf Input:} 
%initial parameter $x_0 \in \R^N$, perturbation constants $\delta_n>0$, step-sizes $\{a_n, b_n\}$, operator $\Upsilon$.
%\For{$n = 0,1,2,\ldots$}	
%	\State Generate $\{d_n^{i}, i=1,\ldots,N\}$, independent of $\{d_m, m=0,1,\ldots,n-1\}$. 
%	\State For any $i=1,\ldots,N$, $d_n^{i}$ is distributed either as an asymmetric Bernoulli (see \eqref{eq:det-proj}) or Uniform $U[-\eta,\eta]$ for some $\eta >0$ (see Remark \ref{remark:unif}). 
%	\PEval
%	    \State Obtain $y_n^+ = f(x_n+\delta_n d_n) + \xi_n^+$.
%  \EndPEval
%	    \PEvalPrime
%	    \State Obtain $y_n^- = f(x_n-\delta_n d_n) + \xi_n^-$.
%	    \EndPEvalPrime
%	    	    \PEvalPrimeDouble
%	    \State Obtain $y_n = f(x_n) + \xi_n$.
%	    \EndPEvalPrimeDouble
%	    \PImpNewton
%		\State Update the parameter and Hessian as follows:
%		\begin{align*}
%		x_{n+1} = & \; x_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(x_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
%where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2rdsa-estimate-ber} and \eqref{eq:psinhat}, respectively. 
%		\EndPImpNewton
%\EndFor
%\State {\bf Return} $x_n.$
%\end{algorithmic}
%\caption{Structure of 2RDSA-IH algorithm.}
%\label{alg:structure}
%\end{algorithm}
%
\begin{algorithm}[!htp]
\KwIn{initial parameter $\theta_0 \in \R^N$, perturbation constants $\delta_n>0$, step-sizes $\{a_n, b_n\}$, operator $\Upsilon$.}
%\KwOut{Policy parameter $\theta$, value function parameter $r$ and feature $\Phi$ }
\begin{enumerate}

 \item \textbf{Execution}: \\
 \For{$n \gets 0,1,2, \ldots, $} {
 \begin{itemize}
 \item Generate $\{d_n^{i}, i=1,\ldots,N\}$, independent of $\{d_m, m=0,1,\ldots,n-1\}$.  
 \item For any $i=1,\ldots,N$, $d_n^{i}$ is distributed either as an asymmetric Bernoulli (see \eqref{eq:det-proj}) or Uniform $U[-\eta,\eta]$ for some $\eta >0$ (see Remark \ref{remark:unif}).
\begin{itemize}
\item \textbf{\emph {Function evaluation 1} }\\ \hspace{4em} Obtain $y_n^+ = f(\theta_n+\delta_n d_n) + \xi_n^+$.
\item \textbf{\emph {Function evaluation 2} }\\ \hspace{4em} Obtain $y_n^- = f(\theta_n-\delta_n d_n) + \xi_n^-$.
\item \textbf{\emph {Function evaluation 3} }\\ \hspace{4em} Obtain $y_n = f(\theta_n) + \xi_n$.
\item \textbf{\emph {Newton step}}\\ \hspace{4.2em} Update the parameter and Hessian as follows:
%\begin{align*}
%		x_{n+1} = & \; x_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(x_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
\begin{align*} 
\theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n).
\end{align*}
where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2rdsa-estimate-ber} and \eqref{eq:psinhat}, respectively. 
\end{itemize}

\end{itemize}
}
\Return{$x_n.$}
\end{enumerate}

\caption{Structure of 2RDSA-IH algorithm.}
\label{alg:structure}
\end{algorithm}

Algorithm \ref{alg:structure} presents the pseudocode and we describe the individual component of 2RDSA-IH below.
\section{Function evaluations}
Let $\delta_n, n\geq 0$ denote a sequence of diminishing positive real numbers and $d_n = (d_n^1,\ldots,d_n^N)\tr$ denote a random perturbation vector at instant $n$,
where the perturbations $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d. and distributed  as follows: 
\begin{equation}
\label{eq:det-proj}
 d_n^i =
  \begin{cases}
   -1 &  \text{ w.p. } \dfrac{(1+\epsilon)}{(2+\epsilon)}, \\
   1+\epsilon &  \text{ w.p. } \dfrac{1}{(2+\epsilon)},
  \end{cases}
\end{equation}
with $\epsilon>0$ being a constant that can be chosen to be arbitrarily small.

The 2RDSA-IH algorithm obtains three function samples $y_n$, $y_n^+$ and $y_n^-$ at $\theta_n$, $\theta_n+\delta_n d_n$ and $\theta_n - \delta_n d_n$, respectively, i.e.,
$y_n = f(\theta_n) + \xi_n$, $y_n^+ = f(\theta_n+\delta_n d_n) + \xi_n^+$ and 
$y_n^- = f(\theta_n-\delta_n d_n) + \xi_n^-$,
where the noise terms $\xi_n, \xi_n^+, \xi_n^-$ satisfy $\E\left[\left.\xi_n^+ + \xi_n^- - 2\xi_n \right| \F_n\right] = 0$ with $\F_n = \sigma(\theta_m,m\le n)$ denoting the underlying sigma-field.

\section{Gradient estimation}
The RDSA estimate of the gradient $\nabla f(\theta_n)$ is given by
\begin{align}
\label{eq:grad-ber}
\widehat\nabla f(\theta_n) = \frac1{1+\epsilon} d_n \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n}\right],
\end{align}
%Note that $E d_n^i = 0$, $E (d_n^i)^2 = 1+\epsilon$ and $E (d_n^i)^4 = \dfrac{(1+\epsilon)(1+(1+\epsilon)^3)}{(2+\epsilon)}$.

\section{Hessian estimation}
\begin{align}
\label{eq:2rdsa-estimate-ber}
&\widehat H_n = M_n \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{\delta_n^2}\right), \text{ where }\\
& M_n =
\left[
\begin{array}{ccc}
\frac{1}{\kappa}\left((d_n^1)^2\!-(1+\epsilon)\right) & \cdots & \frac{1}{2(1+\epsilon)^2}d_n^1 d_n^N\\
\frac{1}{2(1+\epsilon)^2}d_n^2 d_n^1  &  \cdots & \frac{1}{2(1+\epsilon)^2}d_n^2 d_n^N\\
\cdots&\cdots&\cdots\\
\frac{1}{2(1+\epsilon)^2}d_n^N d_n^1 & \cdots &  \frac{1}{\kappa}\left((d_n^N)^2-(1+\epsilon)\right) \\
\end{array}
\right],\nonumber
\end{align}
where $\kappa = \tau \left(1- \dfrac{(1+\epsilon)^2}{\tau}\right)$ and $\tau = E (d_n^i)^4= \dfrac{(1+\epsilon)(1+(1+\epsilon)^3)}{(2+\epsilon)}$, for any $i=1,\ldots,N$. 
%%%%%% Feedback

\section{Improved Hessian estimation}
The Hessian estimate $\widehat H_n$ can be simplified as follows: 
\begin{align}
\widehat H_n & = M_n \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{\delta_n^2}\right) \nonumber \\
& =  M_n \left[\left(\dfrac{f(\theta_n+\delta_n d_n) + f(\theta_n-\delta_n d_n) - 2 f(\theta_n)}{\delta_n^2}\right) \right. +\left. \left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2}\right)\right] \nonumber \\
& = M_n \left(d_n\tr \nabla^2 f(\theta_n) d_n +  O(\delta_n^2) + \left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2} \right)\right).\label{eq:h0}
%& = M_n \left(d_n\tr \nabla^2 f(x_n) d_n +  O(\delta_n^2) + O(\delta_n^{-2})\right). \label{eq:h0}
\end{align}
For the first term on the RHS above, note that 
\begin{align}
&\E[M_n \left(d_n\tr \nabla^2 f(\theta_n) d_n\right) \mid \F_n] =   \E\left[\left. M_n \times\right.\right.\left.\left.\left(\sum\limits_{i=1}^{N-1} (d_n^i)^2 \nabla^2_{ii} f(\theta_n) + 2\sum\limits_{i=1}^N\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n) \right)\right| \F_n\right]. \label{eq:h1}
\end{align}
In analyzing the $l$th diagonal term of the above expression, the following zero-mean term appears (see the proof of Lemma 4 in \cite{prashanth2015rdsa}):
\begin{align}
& \E\left[\left. \left([M_n]_{D}\right)_{l,l} \left(2\sum\limits_{i=1}^{N-1}\sum\limits_{j=i+1}^N d_n^i d_n^j \nabla^2_{ij} f(\theta_n)\right)\right| \F_{n}\right] = 0,\label{eq:diagzero}
\end{align}
where for any matrix $M$, $[M]_{D}$ refers to a matrix that retains only the diagonal entries of $M$ and replaces all the remaining  entries with zero, and $\left([M]_{D}\right)_{i,j}$ refers to the $(i,j)$th entry in $[M]_D$. We shall also use $[M]_{N}$ to refer to a matrix that retains only the off-diagonal entries of $M$, while replaces all the diagonal entries with zero.

The term on the LHS in \eqref{eq:diagzero}, denoted by $\Psi_{n}^{1}(\nabla^2 f(\theta_n))$, can be written in matrix form as follows: 
\begin{align}\label{eq:digmat}
\Psi_{n}^{1}(\nabla^2 f(\theta_n)) = [M_n]_{D}\left(d_{n}\tr \, [\nabla^2 f(\theta_n)]_{N} \, d_{n}\right).
\end{align}

In analyzing the off-diagonal term ($(k,l)$ where $k < l$) of \eqref{eq:h1}, the following zero-mean term appears:
\begin{align}
& \E\left[\left.\left([M_n]_{N}\right)_{k,l}   \left(\sum\limits_{i=1}^N (d_n^i)^2 \nabla^2_{ii} f(\theta_n)\right)\right| \F_n \right] = 0. 
\end{align}
The term on the LHS above, denoted by $\Psi_{n}^{2}(\nabla^2 f(\theta_n))$, can be written in matrix form as follows: 
\begin{align}\label{eq:ndmat}
\Psi_{n}^{2}(\nabla^2 f(\theta_n)) = [M_n]_{N}\left(d_{n}\tr \, [\nabla^2 f(\theta_n)]_{D} \, d_{n}\right).
\end{align}
From the foregoing, the per-iteration Hessian estimate $\widehat H_n$ can be re-written as follows:
\begin{align}
 \E\left[\left.\widehat H_n\right| \F_n\right] = & \nabla^2 f(\theta_n) + \E\left[\left.\Psi_{n}(\nabla^2 f(\theta_n))\right| \F_n\right]  +  O(\delta_n^2)+\E\left[\left.\left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2} \right)\right| \F_n\right], \label{eq:hnhat}
\end{align}
where, for any matrix $H$, 
\begin{align}
\Psi_{n}(H) &= \Psi_{n}^{1}(H) + \Psi_{n}^{2}(H)\nonumber\\
&= [M_n]_{D}\left(d_{n}\tr \, [H]_{N} \, d_{n}\right) +  [M_n]_{N}\left(d_{n}\tr \, [H]_{D} \, d_{n}\right).\label{eq:psi}
\end{align}
In the RHS of \eqref{eq:hnhat}, it is easy to see that  the second term involving $\Psi_{n}$ and the last term involving the noise are zero-mean. Moreover, since the noise is bounded by assumption, the last term in \eqref{eq:hnhat} vanishes asymptotically at the rate $O(\delta_n^{-2})$. So, the error in estimating the Hessian is due to the second term, which involves the perturbations $d_n$. This motivates the term $\widehat \Psi_n$ in the update rule \eqref{eq:e2rdsa}. 
 %and using the fact that $\E\left[\left.\xi_n^+ + \xi_n^- - 2\xi_n \right| \F_n\right] = 0$

Given that we operate in a simulation optimization setting, which implies $\nabla^2 f$ is not known, we construct the feedback term $\widehat \Psi_n$ in \eqref{eq:e2rdsa} by using $\overline H_{n-1}$ as a proxy for $\nabla^2 f$, i.e.,
\begin{align}
\widehat \Psi_n = \Psi_{n} (\overline H_{n-1}).
\label{eq:psinhat}
\end{align}

\section{Step-size optimization}
Unlike the feedback term, adapting the idea of optimizing the step-sizes for 2RDSA is relatively straightforward from the corresponding approach for 2SPSA in \cite{spall-jacobian}. The difference here is that there exists only one $N$-dimensional perturbation vector $d_n$ in our setting, while 2SPSA required two such vectors. This in turn implies that only the perturbation constant $\delta_n$ is needed in optimizing $b_n$.

The optimal choice for $b_n$ in \eqref{eq:2rdsa-H} is the following:
\begin{align}
\label{eq:wieghts}
b_i  = \delta_i^{4}/\sum\limits_{j=0}^{i} \delta_j^{4}.
\end{align}
The main idea behind the above choice is provided below.
From \eqref{eq:hnhat}, we can infer that
\begin{align*}
\E \parallel \widehat H_n \parallel^2 \le \dfrac{C}{\delta_n^4} \text{ for some } C<\infty. 
\end{align*} 
This is because the third term in \eqref{eq:hnhat} vanishes asymptotically, while the fourth term there dominates asymptotically. Moreover, the noise factors in the fourth term in \eqref{eq:hnhat} are bounded above due to (C9) and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.

So, the optimization problem to be solved at instant $n$ is as follows:
\begin{align}
\min_{ \{\tilde b_k\} } \sum \limits_{i=0}^{n} (\tilde b_k)^2 \delta_i^{-4}, \text{ subject to} \label{eq:wn-opt}\\
\tilde b_i \geq 0 ~\forall i \text{ and }\sum \limits_{i=0}^{n} \tilde b_i = 1.
\end{align}
The optimization variable $\tilde b_i$ from the above is related to the Hessian recursion \eqref{eq:2rdsa-H} as follows:
\begin{align}
\label{eq:hess}
\overline H_n = \sum\limits_{i=0}^{n} \tilde b_k(\widehat H_i -\widehat \Psi_i).
\end{align}
The solution to \eqref{eq:wn-opt} is achieved for $\tilde b_i^* = \delta_i^{4}/\sum \limits_{j=0}^{n} \delta_j^{4}, i=1,\ldots,n$. The optimal choice $\tilde b_i^*$ can be translated to the step-sizes $b_i$, leading to \eqref{eq:wieghts}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{remark}\textbf{\textit{(Uniform perturbations)}}
\label{remark:unif}
In \cite{prashanth2015rdsa}, the authors suggest two alternatives for the distribution of random perturbations $d_n$: the asymmetric Bernoulli, which we described earlier and uniform that we outline next.

Choose $d_n^i$, $ i=1,\ldots,N$ to be i.i.d. $U[-\eta,\eta]$ for some $\eta>0$, where $U[-\eta,\eta]$ denotes the uniform distribution on the interval $[-\eta,\eta]$.
Then, the RDSA estimate of the gradient is given by
\begin{align}
\label{eq:grad-unif}
\widehat\nabla f(\theta_n) = \frac3{\eta^2} d_n \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n}\right].
\end{align}

The Hessian estimate in this case is given by
\begin{align}
\label{eq:2rdsa-estimate-unif}
&\widehat H_n =  M_n \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{\delta_n^2}\right), \text{ where } \\
& M_n =
\dfrac{9}{2\eta^4}\left[
\begin{array}{cccc}
\frac{5}{2}\left((d_n^1)^2-\frac{\eta^2}{3}\right) & \cdots & d_n^1 d_n^N\\
d_n^2 d_n^1  &  \cdots & d_n^2 d_n^N\\
\cdots&\cdots&\cdots\\
d_n^N d_n^1 & \cdots &  \frac{5}{2}\left((d_n^N)^2-\frac{\eta^2}{3}\right) \\
\end{array}
\right].\nonumber
\end{align}

The feedback term in \eqref{eq:psinhat} can be easily extended to the case of uniform perturbations by using the $M_n$ as defined above instead of that for the asymmetric Bernoulli case.
\end{remark}
%Henceforth, we shall refer to algorithm \eqref{eq:2rdsa}--\eqref{eq:2rdsa-H} with Hessian estimate \eqref{eq:2rdsa-estimate} as 2RDSA.

\section{Convergence analysis}
\label{sec:2rdsa-results}
We make the same assumptions as those used in the analysis of \cite{prashanth2015rdsa}, with a few minor alterations. The assumptions are listed below:
\begin{enumerate}[label=(\textbf{C\arabic*})]
\item  The function
$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1 i_2 i_3 i_4} f(\theta)$ denotes the $(i_1 i_2 i_3 i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.

%\item  For some $\rho>0$  and almost all $x_n$, the function $f$ is four-times differentiable with a uniformly (in $n$) bounded fourth derivative for all $x$ such that $\left\| x_n - x\right\| \le \rho$.

\item For each $n$ and all $\theta$, there exists a $\rho>0$ not dependent on $n$ and $\theta$, such that $(\theta-\theta^*)\tr \bar f_n(\theta) \ge \rho \left\| \theta_n - \theta\right\|$, where $\bar f_n(\theta) = \Upsilon(\overline H_n)^{-1} \nabla f(\theta)$.

\item $\{\xi_n, \xi_n^+,\xi_n^-, n=1,2,\ldots\}$ are such that, for all $n$, $\E\left[\left. \xi_n^+ + \xi_n^- - 2 \xi_n \right| \F_n\right] = 0$, where $\mathcal{F}_n = \sigma(\theta_m,m\le n)$ denotes the underlying sigma-field..

\item $\{d_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d. and independent of $\F_n$.

\item  The step-sizes $a_n$ and perturbation constants $\delta_n$ are positive, for all $n$ and satisfy
$$\hspace{-1.7em} a_n, \delta_n \rightarrow 0\text{ as } n \rightarrow \infty, 
\sum_n a_n=\infty \text{ and } \sum_n \left(\frac{a_n}{\delta_n}\right)^2 <\infty.$$

\item For each $i=1,\ldots,N$ and any $\rho>0$, 
$P(\{ \bar f_{ni} (\theta_n) \ge 0 \text{ i.o}\} \cap \{ \bar f_{ni} (\theta_n) < 0 \text{ i.o}\} \mid \{ |\theta_{ni} - x^*_i| \ge \rho\quad \forall n\}) =0.$

\item The operator $\Upsilon$ satisfies $\delta_n^2 \Upsilon(H_n)^{-1} \rightarrow 0$ a.s. and  $E(\left\| \Upsilon(H_n)^{-1}\right\|^{2+\zeta}) \le \rho$ for some $\zeta, \rho>0$.

\item For any $\tau >0$ and nonempty $S \subseteq \{1,\ldots,N\}$, there exists a $\rho'(\tau,S)>\tau$ such that 
$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (x-x^*)_i \bar f_{ni}(x)}{\sum_{i \in S} (x-x^*)_i \bar f_{ni}(x)}               \right| < 1 \text{ a.s.}$$
for all $|(x-x^*)_i| < \tau$ when $i \notin S$ and $|(x-x^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
\item For some $\alpha_0, \alpha_1>0$ and for all $n$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n\pm \delta_n d_n)^{2} \le \alpha_1$ and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$. 
\item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
\end{enumerate}
The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions. We remark here that (C1)-(C8) are identical to that in \cite{prashanth2015rdsa}, while (C9) and (C10) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired by \cite{spall-jacobian}.

\begin{lemma}(\textbf{Bias in Hessian estimate})
\label{lemma:2rdsa-bias}
Under (C1)-(C10), with $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-ber}, we have a.s. that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
\begin{align}
\left|\E\left[
\left. \widehat H_n(i,j) \right| \F_n \right] - \nabla^2_{ij} f(\theta_n)\right| = O(\delta_n^2).
\end{align} 
\end{lemma}
\begin{proof}
See Lemma 4 in \cite{prashanth2015rdsa}.
\end{proof}

\begin{theorem}(\textbf{Strong Convergence of Hessian})
\label{thm:2rdsa-H}
Under (C1)-(C10), we have that 
$$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}, respectively, $\widehat H_n$ defined according to \eqref{eq:2rdsa-estimate-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:wieghts}. 
\end{theorem}
\begin{proof}
The first part of the claim regarding $\theta_n$ follows in exactly the same fashion as the proof of Theorem 5 in \cite{prashanth2015rdsa}.  
For proving the claim regarding $\overline H_n$, we closely follow the approach used to prove a corresponding result for 2SPSA (see Theorem 1 in \cite{spall-jacobian}). 
The first step is to prove the following:
\begin{align}
\sum_{k=0}^n \dfrac{\delta_k^4 \left(\widehat H_k - \widehat \Psi_k - \E(\widehat H_k \mid \F_k)\right)}{\sum_{i=0}^n \delta_i^4} \rightarrow 0.
\label{eq:step1}
\end{align}

By a completely parallel argument to that used in the proof of Theorem 1 in \cite{spall-jacobian}, we obtain: For any $i,j = 1,\ldots,N$,
\begin{align*}
\E \left[\left((\widehat {H}_k)_{i,j} - (\widehat {\Psi}_k)_{i,j} - \E((\widehat {H}_k)_{i,j} \mid \F_k)\right)^2\right] = O(\delta_k^{-4}).
\end{align*}
Now \eqref{eq:step1} follows by an application of Kronecker's Lemma along with the martingale convergence theorem (see Theorem 6.2.1 of \cite{lahaprobability}).

From Lemma \ref{lemma:2rdsa-bias}, we have 
$$ \E[ \widehat H_k \mid \F_k] = \nabla^2 f(\theta_n) + O(\delta_n^2) \text{ a.s.}$$
Since the Hessian is continuous near $\theta_n$ and $\theta_n$ converges almost surely to $\theta^*$, we have
% \footnote{Since assumptions (C1)-(C10) here are similar to that in \cite{prashanth2015rdsa}, Theorem 5 of \cite{prashanth2015rdsa} holds here as well. In particular, this implies almost sure convergence of $x_n$ to $x^*$.}
\begin{align*}
\sum_{k=0}^n \dfrac{\delta_k^4 \left(\E(\widehat H_k \mid \F_k)\right)}{\sum_{i=0}^n \delta_i^4} 
=&\sum_{k=0}^n \dfrac{\delta_k^4 \left(\nabla^2 f(\theta_n) + O(\delta_n^2)\right)}{\sum_{i=0}^n \delta_i^4}\\
=&\sum_{k=0}^n \dfrac{\delta_k^4 \left(\nabla^2 f(x^*) + o(1)\right)}{\sum_{i=0}^n \delta_i^4}\\
&\rightarrow \nabla^2 f(x^*) \text{ a.s. as } n \rightarrow \infty.
\end{align*}
The last step above follows from Toeplitz Lemma (see p. 89 of \cite{lahaprobability}) after observing that $\sum_{i=0}^n \delta_i^4 \rightarrow \infty$ due to (C10). 
The main claim now follows since 
$$ \overline H_n = \sum_{k=0}^n \dfrac{\delta_k^4 \left(\widehat H_k - \Psi_k \right)}{\sum_{i=0}^n \delta_i^4}.$$
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%%% Quadratic case
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
%\begin{enumerate}[label=(\textbf{C\arabic*}),resume]
%\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
%\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
%\end{enumerate}
%\begin{theorem}(\textbf{Quadratic case - Convergence rate})
%\label{thm:quad-bound}
%Assume (C4), (C10), (C11) and (C12) and also that the setting is noise-free. 
%Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
%\begin{align}
%\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
%\label{eq:quad-bigo}
%\end{align}
%\end{theorem}
%\begin{proof}
%Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:hnhat} as follows:
%\begin{align}
% \widehat H_n =    \Phi_n(\nabla^2 f(\theta_n)) &+\Psi_{n}(\nabla^2 f(\theta_n)),\label{eq:hnhat-ext}
%\end{align}
%where $\Psi_n(H)$ is defined in \eqref{eq:psi} and for any matrix $H$, 
%$$\Phi_{n}(H) = [M_n]_{D}\left(d_{n}\tr \, [H]_{D} \, d_{n}\right) +  [M_n]_{N}\left(d_{n}\tr \, [H]_{N} \, d_{n}\right).$$
%The proof involves the following steps:
%\begin{description}
%  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
%  \item[Step 2:] We unroll the recursion \eqref{eq:2rdsa-H} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
%  \item[Step 3:] We derive the main result in \eqref{eq:quad-bigo} using a proof by contradiction. 
%\end{description}
%\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
%This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
%\noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
% From \eqref{eq:2rdsa-H} and \eqref{eq:hnhat-ext},  we have
% \begin{align}
% \Lambda_n &= \Lambda_{n-1} - b_n ( \overline H_{n-1} - \hat H_n + \hat \Psi_n) \nonumber\\
%&= (1-b_n) \Lambda_{n-1} - b_n (H^* + \hat \Psi_n - \hat H_n ) \nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n (H^*+\hat \Psi_n -\Phi_n(H^*) - \Psi_n(H^*)\nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n \Psi_n(\Upsilon(\overline H_{n-1})) + b_n (\Phi_n(H^*) - H^*)\nonumber\\
%&= (1-b_n) \Lambda_{n-1} - b_n\Psi_n(\Lambda_{n-1}')+ b_n (\Phi_n(H^*) - H^*),\label{eq:step1last}
% \end{align}
%where $\Lambda_n' = \Upsilon(\overline H_n) - H^*$.
%The equality in \eqref{eq:step1last} follows from \eqref{eq:psi}. 
%Unrolling the recursion \eqref{eq:step1last}, we obtain 
% \begin{align}\label{lambda-exp}
%  \Lambda_n  = \left[ \prod_{k=1}^n (1-b_k) \right]\Lambda_0 
%&- \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k \Psi_k(\Lambda_{k-1}') \nonumber\\
%&+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k (\Phi_k(H^*) - H^*) \hspace{0.2cm} a.s.
% \end{align}
% %%(note : $\prod_{j=n+1}^n (1-b_j) = 1$ for all n).
% %%Let us  characterise $trace[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{lambda-exp}. From independence of $d_k$ along k, \eqref{lambda-exp} represent martingale difference sequence, leading to 
%Squaring on both sides and taking expectations, we obtain
% \begin{align}
%  \E (\Lambda_n \tr \Lambda_n) =  &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \E (\Lambda_0 \tr \Lambda_0) + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \E (\Psi_k(\Lambda_{k-1}') \tr \Psi_k(\Lambda_{k-1}'))  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \times \E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*)) \nonumber\\
%&-2 \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2 \E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right).\label{eq:exp-lambda2}
% \end{align}
%The equality above uses the fact that $\E(\Psi_k(\Lambda_{k-1}'))=0$, $\E(\Phi_k(H^*) - H^*)=0$, which get rid of the corresponding cross terms with the first term on RHS of \eqref{lambda-exp}. Notice that there are only $n$ terms in the last term in the RHS of \eqref{eq:exp-lambda2} because $d_k$ are identically distributed for each $k$ and across $k$.  
%% %
%We now characterize $\text{trace}[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{eq:exp-lambda2} as follows:
%%From the independence of $d_k$'s and $\Lambda_{k-1}'$ and using Cauchy-Schwartz inequality we get 
% \begin{align}\label{eq:trace}
% \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]  \nonumber\\ &- 2 \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right].
% \end{align}
% % \begin{align}\label{eq:trace}
%% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq \left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] + \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
%% \end{align}
%
%As in the proof of Theorem 3 of \cite{spall-jacobian}, observe that $1-b_k = e^{-b_k}(1-O(b_k^2))$ and since $0 < b_k <1$, we have that the $O(b_k^2)$ term is strictly positive. 
%Letting $\Gamma_{ij} = \sum_{k=i}^j b_k$ with $\Gamma_{nn} = 1$ and $\beta_{kn} = \left[\prod_{i=k+1}^n (1- O(b_i^2))\right]^2$, we can simplify \eqref{eq:trace} as follows:
%\begin{align}\label{eq:wij}
% \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &- 2 e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2   \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right] .
% \end{align} 
%% \begin{align}\label{eq:wij}
%% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
%% \end{align} 
%Comparing the sum with integrals, we obtain
%\begin{align*}
%\Gamma_{ij} = \int_i^j \frac{b_0}{x^r} dx + O(1) &= \left(\frac{b_0}{1-r}\right)(j^{1-r}-i^{1-r}) +O(1),
%\end{align*}
%where we have used the facts that $0 < b_k < 1, \forall k \ge 2$ and 
%$\sum_{k=i}^j b_k \to \infty$ as $j-i \to \infty$ since 
%$b_k = b_0/k^r$ with $r > 0.5$.
%Observing that $\beta_{kn}$ are uniformly upper-bounded, say by $\bar \beta_n$, we have
%\begin{align}\label{eq:thmwts}
% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &- 2 \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}   \times \text{trace} \left[\E\left(\Psi_k(\Lambda_{k-1}) \tr (\Phi_k(H^*) - H^*)\right)\right].
%\end{align} 
%\begin{align}\label{eq:thmcomb}
%& \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \text{trace}\left[\E [((\Phi_k(H^*) - H^*) - 2 \Psi_k(\Lambda_{k-1}))\tr  (\Phi_k(H^*) - H^*)]\right] .
%\end{align} 
%%
%%\begin{align}\label{eq:thmwts}
%% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
%%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \nonumber\\
%%& \hspace{3.5cm} \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
%%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\
%%&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\&\hspace{3.5cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2}\nonumber\\&\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
%%\end{align}
%\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
%In comparison to Eq. (7.6) in \cite{spall-jacobian} that corresponds to \eqref{eq:thmwts} for the N-SPSA-4 setting, there are two extra terms - the third and the fourth - in the RHS of \eqref{eq:thmwts}. 
%Consider the third term on the RHS of \eqref{eq:thmwts}. As a consequence of (C9) and the construction of random perturbations $d_k$, it can be seen that $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$ is uniformly bounded above, say by $\top$, independent of $k$. Thus, the third term is equivalent to the following:
%\begin{align}
% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \top.
%\label{eq:term3}
%\end{align}
%Now by moving out side exponential term to inside and changing the summation to integration, we can rewrite the above equation as follows
%\begin{align}
%  \overline{\overline{c_n}}  \int_1^n  \frac{1}{k^{2 r}} .
%\label{eq:term3}
%\end{align}
%Where $\overline{\overline{c_n}}$ is a constant. By above equation we can conclude that the third term is of $O(\frac{1}{k^{2 r -1}})$
%%Now, observe that $\lim_{n\rightarrow \infty} \sum_{k=1}^n \frac{1}{k^{2 r}} < \infty$ as $1/2 < r < 1$ and $e^{2 b_0 k^{1-r}/(1-r)}$ diverges as $n\rightarrow \infty$ and hence, by Kronecker's Lemma, the term in \eqref{eq:term3} vanishes asymptotically. 
%%
%%Now consider the fourth term. By using the uniform upper-bound $\top$ on $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$, we can rewrite  the fourth term in \eqref{eq:thmwts} as follows: 
%% \begin{align}\label{eq:fourthterm}
%% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \sqrt{\top} \nonumber \\ \hspace{3cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2},
%% \end{align}
%%where, as in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a linear fashion and then returns the trace of the resulting $N\times N$ matrix.
%
%\textbf{Part (iii) needed fix} 
%%of Theorem 3 in \cite{spall-jacobian} uses a proof by contradiction to show that the first and second terms in \eqref{eq:thmwts} are both of the order $O(e^{-2b_0n^{1-r} / {1-r}})$. By a completely parallel argument to that used for the second term in \cite{spall-jacobian}, it can be seen that the fourth term in our setting vanishes at a faster rate than the second term.  The claim follows. 
%\end{proof}