### 2017-01-17 second-order-spsa-3.tex
counter conditions problem resoved

 \chapter{Second-order SPSA-3 with improved hessian estimation (2SPSA-3-IH)}
\label{sec:2spsa-3-ih}
The second-order SPSA-3 with improved Hessian estimate performs an update iteration similar to that of 2RDSA-IH proposed in chapter \ref{sec:2rdsa-ih} (See \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}). The main difference in the iterates is that 2SPSA-3-IH uses gradient and Hessian estimates according to \eqref{eq:grad-spsa} and \eqref{eq:2spsa-ber}. In this chapter we present the improvements to 2SPSA-3 Hessian recursion \eqref{eq:2spsa-3-ih} by incorporating a zero-mean feedback term and a general step-size. Note that though these ideas of improving the Hessian recursion are similar to the ideas presented in chapter \ref{sec:2rdsa-ih} for 2RDSA-IH, the derivation of the feedback term does not directly follow. We show that the proposed improvements to Hessian estimation in 2SPSA-3 are such  that the resulting 2SPSA-3-IH algorithm is provably convergent, in particular, the Hessian estimate $\overline H_n$ of 2SPSA-3-IH converges almost surely to the true Hessian. Moreover, we show, for the special case of a quadratic objective in noise-free setting, that the 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. The advantage with 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.

%The recursion \eqref{eq:e2rdsa} is identical to that in 2RDSA, while the Hessian estimation recursion \eqref{eq:2rdsa-H} differs as follows:\\
%\begin{inparaenum}[\bfseries (i)]
%\item  $\widehat \Psi_n$ is a zero-mean feedback term that reduces the error in Hessian estimate; and\\
%\item $b_n$ is a general step-size that we optimize to improve the Hessian estimate.
%\end{inparaenum}
%
%On the other hand, $\widehat H_n$ is identical to that in 2RDSA, i.e., it estimates the true Hessian in each iteration using $3$ function evaluations. For the sake of completeness, we provide below the construction for $\widehat\nabla f(\theta_n)$ and $\widehat H_n$ using asymmetric Bernoulli perturbations, before we present the feedback term that reduces the error in $\widehat H_n$.
\begin{algorithm}[!htp]
\KwIn{initial parameter $\theta_0 \in \R^N$, perturbation constants $\delta_n>0$, step-sizes $\{a_n, b_n\}$, operator $\Upsilon$.}

\begin{enumerate}

 \item \textbf{Execution}: \\
 \For{$n \gets 0,1,2, \ldots, $} {
 \begin{itemize}
 \item Generate $\{\Delta_n^{i}, i=1,\ldots,N\}$,$\{\widehat\Delta_n^{i}, i=1,\ldots,N\}$ independent of $\{\Delta_m,\widehat\Delta_m, m=0,1,\ldots,n-1\}$.  
 \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (C17) of section \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p. $1/2$).
\begin{itemize}
\item \textbf{\emph {Function evaluation 1} }\\ \hspace{4em} Obtain $y_n^+ = f(\theta_n+\delta_n \Delta_n +\delta_n \widehat\Delta_n) + \xi_n^+$.
\item \textbf{\emph {Function evaluation 2} }\\ \hspace{4em} Obtain $y_n^- = f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) + \xi_n^-$.
\item \textbf{\emph {Function evaluation 3} }\\ \hspace{4em} Obtain $y_n = f(\theta_n) + \xi_n$.
\item \textbf{\emph {Newton step}}\\ \hspace{4.2em} Update the parameter and Hessian as follows:
%\begin{align*}
%		\theta_{n+1} = & \; \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
\begin{align} \label{eq:2spsa-3-theta}
\theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n).\label{eq:2spsa-3-ih}
\end{align}
where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2spsa-ber} and \eqref{eq:psinhatspsa3}, respectively. 
\end{itemize}

\end{itemize}
}
\Return{$\theta_n.$}
\end{enumerate}

\caption{Structure of 2SPSA-3-IH algorithm.}
\label{alg:spsa-3-ih-structure}
\end{algorithm}

Algorithm \ref{alg:spsa-3-ih-structure} presents the pseudocode and we describe the individual components of 2SPSA-3-IH below.
\section{Function evaluations}
Let $y_n$, $y_n^+$ and $y_n^-$ denote the function evaluations at $\theta_n$, $\theta_n+\delta_n \Delta_n+\delta_n \widehat \Delta_n$ and $\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n$ respectively, i.e., 
$y_n = f(\theta_n) + \xi_n$, $y_n^+ = f(\theta_n+\delta_n \Delta_n +\delta_n \widehat\Delta_n) + \xi_n^+$ and 
$y_n^- = f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) + \xi_n^-$,
where the noise terms $\xi_n, \xi_n^+, \xi_n^-$ satisfy $\E\left[\left.\xi_n^+ + \xi_n^- - 2\xi_n \right| \F_n, \Delta_n, \widehat \Delta_n\right] = 0$ with $\F_n = \sigma(\theta_m,m <  n)$ denoting the underlying sigma-field.

Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and $\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$ , $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vector at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ are independent identically distributed (i.i.d), mean-zero random varibles having finite inverse movements of order greater than 2 and satisfy the condition (C17) of section \ref{sec:2spsa-results}.

\section{Gradient estimate}
The SPSA estimate of the gradient $\nabla f(\theta_n)$ is given by
\begin{align}
\label{eq:grad-spsa}
\widehat\nabla_{(i)} f(\theta_n) =  \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n \Delta_n^{(i)}}\right],
\end{align}
where the perturbations $\Delta_n^i$, $i=1,\ldots,N$ are i.i.d. and as mentioned above.

%Note that $E d_n^i = 0$, $E (d_n^i)^2 = 1+\epsilon$ and $E (d_n^i)^4 = \dfrac{(1+\epsilon)(1+(1+\epsilon)^3)}{(2+\epsilon)}$.

\section{Hessian estimate}
The $i,j$th entry of the Hessian estimate in this case is given by
\begin{align}
\label{eq:2spsa-ber}
&\left(\widehat H_n\right)_{ij} =  \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right), 
\end{align}
%%%%%% Feedback
\section{Feedback term $\widehat \Psi_n$}
The $i,j$th term of the Hessian estimate $\widehat H_n$ can be simplified as follows: 
\begin{align}
\left(\widehat H_n \right)_{ij} & =  \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right) \nonumber \\
%& =  \left[\left(\dfrac{f(\theta_n+\delta_n \Delta_n+\delta_n \widehat \Delta_n) + f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) - 2 f(\theta_n)}{\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right) \right.\nonumber \\&\hspace{10em}+\left. \left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2}\right)\right] \nonumber \\
& = \left( \frac{(\Delta_n+ \widehat \Delta_n)\tr \nabla^2 f(\theta_n) (\Delta_n+ \widehat \Delta_n)}{2\Delta_n^{(i)} \widehat \Delta_n^{(j)}} +  O(\delta_n^2) + \right.\left.\left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}} \right)\right).\label{eq:h3}
%& = M_n \left(d_n\tr \nabla^2 f(\theta_n) d_n +  O(\delta_n^2) + O(\delta_n^{-2})\right). \label{eq:h0}
\end{align}
For the first term on the RHS above, note that
\begin{align}
\E\left[\left.  \frac{(\Delta_n+ \widehat \Delta_n)\tr \nabla^2 f(\theta_n) (\Delta_n+ \widehat \Delta_n)}{2\Delta_n^{(i)} \widehat \Delta_n^{(j)}}  \right | \F_n\right] &=\E\left[\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \Delta_n^m}{2\Delta_n^i \widehat\Delta_n^j} + \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j}  \right.\nonumber \\&\hspace{2.5cm}+\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \left. \frac{\widehat\Delta_n^l \nabla^2_{lm} f(\theta_n)\widehat \Delta_n^m}{2\Delta_n^i \widehat\Delta_n^j}\right| \F_n\right].\label{eq:2spsaexp}
\end{align}
In analysing the RHS of the above expression, the first an the third terms are mean-zero (See Proposition 4.2 in \cite{bhatnagar2015simultaneous}):
\begin{align}
\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = 0 \hspace{0.2cm} and \hspace{0.2cm}\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\widehat\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j }\right | \F_n\right] = 0.
\end{align}
%\begin{align}
%\E\left[\left. \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = \nabla^2_{ij} f(\theta_n) \hspace{0.3cm}a.s..
%\end{align}
From the above equations the  terms that are in expectation, denoted by  $\Psi_n^1(\nabla^2 f(\theta_n))$, can be written in matrix form as follows:
\begin{align}
\Psi_n^1(\nabla^2 f(\theta_n)) =\frac{1}{2} M_n \left[\Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + \widehat\Delta_n\tr \nabla^2 f(\theta_n) \widehat\Delta_n\right].
\end{align}
where $M_n = [1/\Delta_n^1, \ldots ,1/\Delta_n^N]\tr [1/\Delta_n^1, \ldots ,1/\Delta_n^N]$.
Now consider the second term in RHS of the \eqref{eq:2spsaexp}. It can be re-written as follows:
\begin{align}
\E\left[\left. \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = \E\left[\left.\nabla^2_{ij} f(\theta_n) + \frac{1}{\Delta_n^i \widehat\Delta_n^j} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m \right | \F_n\right]
\end{align}
In analysing the RHS of the above expression, the second term in the expectation  is mean-zero term:  
\begin{align}
\E\left[\left. \frac{1}{\Delta_n^i \widehat\Delta_n^j} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m \right | \F_n\right] = 0.
\end{align}
The term on the LHS above, can be denoted by $\Psi_n^2(\nabla^2 f(\theta_n))$, can be written in the matrix form as follows:
\begin{align}
\Psi_n^2(\nabla^2 f(\theta_n)) = \widehat N_n\tr \nabla^2 f(\theta_n) N_n  + \widehat N_n\tr \nabla^2 f(\theta_n) + \nabla^2 f(\theta_n) N_n.
\end{align}
where $N_n,\widehat N_n$ are defined as follows $N_n = \Delta_n \left[\frac{1}{\Delta_n^1},\ldots,\frac{1}{\Delta_n^N}\right] - I_N$ and $\widehat N_n = \widehat \Delta_n \left [\frac{1}{\widehat\Delta_n^1},\ldots,\frac{1}{\widehat \Delta_n^N}\right] - I_N$ and $I_N$ is  identity matrix of size $N \times N$.
From the foregoing, the per-iteration Hessian estimate $\widehat H_n$ can be re-written as follows:
\begin{align}
\widehat H_n = & \nabla^2 f(\theta_n) + \Psi_{n}(\nabla^2 f(\theta_n)) +  O(\delta_n^2) + O(\delta_n^{-2})  \label{eq:spsahnhat}
\end{align}
where, for any matrix $H$, 
\begin{align}
\Psi_{n}(H) &=  \Psi_{n}^{1}(H) + \Psi_{n}^{2}(H)\nonumber\\
&=  \frac{1}{2}M_n \left[\Delta_n\tr H \Delta_n + \widehat\Delta_n\tr H \widehat\Delta_n\right] + \widehat N_n\tr H N_n + \widehat N_n\tr H + H N_n.\label{eq:spsapsi}
\end{align}
Given that we operate in a simulation optimization setting, which implies $\nabla^2 f$ is not known, we construct the feedback term $\widehat \Psi_n$ in \eqref{eq:2spsa-3-ih} by using $\overline H_{n-1}$ as a proxy for $\nabla^2 f$, i.e.,
\begin{align}
\widehat \Psi_n = \Psi_{n} (\overline H_{n-1}).
\label{eq:psinhatspsa3}
\end{align}

\paragraph{\textbf{Optimizing the step-sizes }$b_n$}

The optimal choice for $b_n$ in \eqref{eq:2spsa-3-ih}  is the following:
\begin{align}
\label{eq:spsawieghts}
b_i  = \delta_i^{4}/\sum\limits_{j=0}^{i} \delta_j^{4}.
\end{align}
The main idea behind the above choice is provided below.
From \eqref{eq:spsahnhat}, we can infer that
\begin{align*}
\E \parallel \widehat H_n \parallel^2 \le \dfrac{C}{\delta_n^4} \text{ for some } C<\infty. 
\end{align*} 
This is because the third term in \eqref{eq:spsahnhat} vanishes asymptotically, while the fourth term there dominates asymptotically. Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (C22)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.

So, the optimization problem to be solved at instant $n$ is as follows:
\begin{align}
\sum \limits_{i=0}^{n} (\tilde b_k)^2 \delta_i^{-4}, \text{ subject to} \label{eq:spsawn-opt}\\
\tilde b_i \geq 0 ~\forall i \text{ and }\sum \limits_{i=0}^{n} \tilde b_i = 1.
\end{align}
The optimization variable $\tilde b_i$ from the above is related to the Hessian recursion \eqref{eq:2spsa-3-ih} as follows:
\begin{align}
\label{eq:spsahess}
\overline H_n = \sum\limits_{i=0}^{n} \tilde b_k(\widehat H_i -\widehat \Psi_i).
\end{align}
The solution to \eqref{eq:spsawn-opt} is achieved for $\tilde b_i^* = \delta_i^{4}/\sum \limits_{j=0}^{n} \delta_j^{4}, i=1,\ldots,n$. The optimal choice $\tilde b_i^*$ can be translated to the step-sizes $b_i$, leading to \eqref{eq:spsawieghts}.

\section{Convergence analysis for 2SPSA-3-IH}
\label{sec:2spsa-results}
We make the same assumptions as those used in the analysis of \cite{prashanth2015rdsa}, with a few minor alterations. The assumptions are listed below:
\begin{pvn}
\item  The function
$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1 i_2 i_3 i_4} f(\theta)$ denotes the $(i_1 i_2 i_3 i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.

%\item  For some $\rho>0$  and almost all $\theta_n$, the function $f$ is four-times differentiable with a uniformly (in $n$) bounded fourth derivative for all $x$ such that $\left\| \theta_n - x\right\| \le \rho$.

\item For each $n$ and all $x$, there exists a $\rho>0$ not dependent on $n$ and $\theta$, such that $(\theta-\theta^*)\tr \bar f_n(\theta) \ge \rho \left\| \theta_n - \theta\right\|$, where $\bar f_n(\theta) = \Upsilon(\overline H_n)^{-1} \nabla f(\theta)$.

\item $\{\xi_n, \xi_n^+,\xi_n^-, n=1,2,\ldots\}$ are such that, for all $n$, $\E\left[\left. \xi_n^+ + \xi_n^- - 2 \xi_n \right| \F_n, \Delta_n, \widehat \Delta_n \right] = 0$, where $\mathcal{F}_n = \sigma(\theta_m,m < n)$ denotes the underlying sigma-field..

\item $\{\Delta_n^i,\widehat\Delta_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d independent of $\F_n$ and for some $\alpha_0 > 0$ and for all $n,l$ $|\Delta_{nl}|\le \alpha_0$, $\Delta_{nl}$ is symmetrically distributed about 0, $\Delta_{nl}$ are mutually independent across $n$ and $l$ and they satisfy $\E (\Delta_{nl}^{-2})$,$\E (\widehat\Delta_{nl}^{-2}) \le \alpha_0$.

\item  The step-sizes $a_n$ and perturbation constants $\delta_n$ are positive, for all $n$ and satisfy
$$\hspace{-1.7em} a_n, \delta_n \rightarrow 0\text{ as } n \rightarrow \infty, 
\sum_n a_n=\infty \text{ and } \sum_n \left(\frac{a_n}{\delta_n}\right)^2 <\infty.$$

\item For each $i=1,\ldots,N$ and any $\rho>0$, 
$P(\{ \bar f_{ni} (\theta_n) \ge 0 \text{ i.o}\} \cap \{ \bar f_{ni} (\theta_n) < 0 \text{ i.o}\} \mid \{ |\theta_{ni} - \theta^*_i| \ge \rho\quad \forall n\}) =0.$

\item The operator $\Upsilon$ satisfies $\delta_n^2 \Upsilon(H_n)^{-1} \rightarrow 0$ a.s. and  $E(\left\| \Upsilon(H_n)^{-1}\right\|^{2+\zeta}) \le \rho$ for some $\zeta, \rho>0$.

\item For any $\tau >0$ and nonempty $S \subseteq \{1,\ldots,N\}$, there exists a $\rho'(\tau,S)>\tau$ such that 
$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}{\sum_{i \in S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}               \right| < 1 \text{ a.s.}$$
for all $|(\theta-\theta^*)_i| < \tau$ when $i \notin S$ and $|(\theta-\theta^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
\item For some $\alpha_0, \alpha_1, \alpha_2 >0$ and for all $n,l,m$, $\E {\xi_n}^{2} \le \alpha_0$, $\E {\xi_n^{\pm}}^{2} \le \alpha_0$, $\E f(\theta_n)^{2} \le \alpha_1$,  $\E f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2},\E f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2} \le \alpha_1$, 
%$\E \left[ f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[| f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
%$\E \left[ f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[ |f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_2} \mid \F_n\right]$,
 $\E\left[\left. (\xi_n^+ + \xi_n^- - 2 \xi_n)^2/(\Delta_{nl}\widehat\Delta_{nm})^2 \right| \F_n \right] \le \alpha_1$ 
and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_1$. 
\item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
\end{pvn}
The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions. We remark here that (C14)-(C21) are identical to that in \cite{prashanth2015rdsa}, while (C22) and (C23) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n \right\|^2$ and $\delta_n$, respectively and these are inspired by \cite{spall-jacobian}.

\begin{lemma}(\textbf{2SPSA-3-IH Bias in Hessian estimate})
\label{lemma:2spsa-bias}
Under (C14)-(C23), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s. that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{ij}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
\begin{align}
\left|\E\left[
\left. \widehat H_n(i,j) \right| \F_n \right] - \nabla^2_{ij} f(\theta_n)\right| = O(\delta_n^2).
\end{align} 
\end{lemma}
\begin{proof}
See Proposition 4.2 in \cite{bhatnagar2015simultaneous}.
\end{proof}

\begin{theorem}(\textbf{2SPSA-3-IH Strong Convergence of Hessian})
\label{thm:2spsa-H}
Under (C14)-(C23), we have that 
$$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:2spsa-3-theta} and \eqref{eq:2spsa-3-ih}, respectively, $\widehat H_n$ defined according to \eqref{eq:2spsa-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:spsawieghts}. 
\end{theorem}
\begin{proof}
This proof is similar to the proof of the theorem \ref{thm:2rdsa-H}
\end{proof}

We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
\begin{pvn}
\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and $\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
\end{pvn}
\begin{theorem}(\textbf{2SPSA-3-IH Quadratic case - Convergence rate})
\label{thm:spsaquad-bound}
Assume (C17), (C23), (C24) and (C25) and also that the setting is noise-free. 
Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
\begin{align}
\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
\label{eq:spsaquad-bigo}
\end{align}
\end{theorem}
\begin{proof}
Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:spsahnhat} as follows:
\begin{align}
 \widehat H_n =    \nabla^2 f(\theta_n) +\Psi_{n}(\nabla^2 f(\theta_n)),\label{eq:spsahnhat-ext}
\end{align}
where $\Psi_n(H)$ is defined in \eqref{eq:psinhatspsa3}.
The proof involves the following steps:
\begin{description}
  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
  \item[Step 2:] We unroll the recursion \eqref{eq:2spsa-3-ih} and then derive convenient representation for $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
  \item[Step 3:] We derive the main result in \eqref{eq:spsaquad-bigo} using a proof by contradiction. 
\end{description}
\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
\noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
This part in exactly the same manner as part (ii) in Theorem 3 of \cite{spall-jacobian}.\\
\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
This part in exactly the same manner as part (iii) in Theorem 3 of \cite{spall-jacobian}.\\
\end{proof}