### 2017-01-23 second-order-spsa-3.tex

 \chapter{Second-order SPSA-3 with improved hessian estimation (2SPSA-3-IH)}
\label{sec:2spsa-3-ih}
The second-order SPSA-3 with improved Hessian estimate performs an update iteration similar to that of 2RDSA-IH proposed in chapter \ref{sec:2rdsa-ih} (See \eqref{eq:e2rdsa} and \eqref{eq:2rdsa-H}). The main difference in the iterates is that 2SPSA-3-IH uses gradient and Hessian estimates according to \eqref{eq:grad-spsa} and \eqref{eq:2spsa-ber}. In this chapter we present the improvements to 2SPSA-3 Hessian recursion \eqref{eq:2spsa-3-ih} by incorporating a zero-mean feedback term and a general step-size. Note that though these ideas of improving the Hessian recursion are similar to the ideas presented in chapter \ref{sec:2rdsa-ih} for 2RDSA-IH, the derivation of the feedback term does not directly follow. We show that the proposed improvements to Hessian estimation in 2SPSA-3 are such  that the resulting 2SPSA-3-IH algorithm is provably convergent, in particular, the Hessian estimate $\overline H_n$ of 2SPSA-3-IH converges almost surely to the true Hessian. Moreover, we show, for the special case of a quadratic objective in noise-free setting, that the 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. The advantage with 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.

%The recursion \eqref{eq:e2rdsa} is identical to that in 2RDSA, while the Hessian estimation recursion \eqref{eq:2rdsa-H} differs as follows:\\
%\begin{inparaenum}[\bfseries (i)]
%\item  $\widehat \Psi_n$ is a zero-mean feedback term that reduces the error in Hessian estimate; and\\
%\item $b_n$ is a general step-size that we optimize to improve the Hessian estimate.
%\end{inparaenum}
%
%On the other hand, $\widehat H_n$ is identical to that in 2RDSA, i.e., it estimates the true Hessian in each iteration using $3$ function evaluations. For the sake of completeness, we provide below the construction for $\widehat\nabla f(\theta_n)$ and $\widehat H_n$ using asymmetric Bernoulli perturbations, before we present the feedback term that reduces the error in $\widehat H_n$.
\begin{algorithm}[!htp]
\KwIn{initial parameter $\theta_0 \in \R^N$, perturbation constants $\delta_n>0$, step-sizes $\{a_n, b_n\}$, operator $\Upsilon$.}

\begin{enumerate}

 \item \textbf{Execution}: \\
 \For{$n \gets 0,1,2, \ldots, $} {
 \begin{itemize}
 \item Generate $\{\Delta_n^{i}, i=1,\ldots,N\}$,$\{\widehat\Delta_n^{i}, i=1,\ldots,N\}$ independent of $\{\Delta_m,\widehat\Delta_m, m=0,1,\ldots,n-1\}$.  
 \item For any $i=1,\ldots,N$, $\Delta_n^{i}$,$\widehat\Delta_n^{i}$  satisfy condition (A23) of chapter \ref{sec:2spsa-results},(most popular distribution being used is symmetric Bernoulli distribution, $\pm 1$ w.p. $1/2$).
\begin{itemize}
\item \textbf{\emph {Function evaluation 1} }\\ \hspace{4em} Obtain $y_n^+ = f(\theta_n+\delta_n \Delta_n +\delta_n \widehat\Delta_n) + \xi_n^+$.
\item \textbf{\emph {Function evaluation 2} }\\ \hspace{4em} Obtain $y_n^- = f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) + \xi_n^-$.
\item \textbf{\emph {Function evaluation 3} }\\ \hspace{4em} Obtain $y_n = f(\theta_n) + \xi_n$.
\item \textbf{\emph {Newton step}}\\ \hspace{4.2em} Update the parameter and Hessian as follows:
%\begin{align*}
%		\theta_{n+1} = & \; \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n), \\
%\overline H_n = &\; (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),
%\end{align*}
\begin{align} \label{eq:2spsa-3-theta}
\theta_{n+1} = \,& \theta_n - a_n \Upsilon(\overline H_n)^{-1}\widehat\nabla f(\theta_n),\\
\overline H_n =\,& (1-b_{n})  \overline H_{n-1} + b_{n} ( \widehat H_n - \widehat \Psi_n),\label{eq:2spsa-3-ih}
\end{align}
where $\widehat H_n$ and $\widehat \Psi_n$ are chosen according to \eqref{eq:2spsa-ber} and \eqref{eq:psinhatspsa3}, respectively. 
\end{itemize}

\end{itemize}
}
\Return{$\theta_n.$}
\end{enumerate}

\caption{Structure of 2SPSA-3-IH algorithm.}
\label{alg:spsa-3-ih-structure}
\end{algorithm}

Algorithm \ref{alg:spsa-3-ih-structure} presents the pseudocode and we describe the individual components of 2SPSA-3-IH below.
\section{Function evaluations}
Let $y_n$, $y_n^+$ and $y_n^-$ denote the function evaluations at $\theta_n$, $\theta_n+\delta_n \Delta_n+\delta_n \widehat \Delta_n$ and $\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n$ respectively, i.e., 
$y_n = f(\theta_n) + \xi_n$, $y_n^+ = f(\theta_n+\delta_n \Delta_n +\delta_n \widehat\Delta_n) + \xi_n^+$ and 
$y_n^- = f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) + \xi_n^-$,
where the noise terms $\xi_n, \xi_n^+, \xi_n^-$ satisfy $\E\left[\left.\xi_n^+ + \xi_n^- - 2\xi_n \right| \F_n, \Delta_n, \widehat \Delta_n\right] = 0$ with $\F_n = \sigma(\theta_m,m <  n)$ denoting the underlying sigma-field.

Further, $\delta_n, n\geq 0$ is a sequence of diminishing positive real numbers and 
$\Delta_n = (\Delta_n^1,\ldots,\Delta_n^N)\tr$, $\widehat\Delta_n = (\widehat\Delta_n^1,\ldots,\widehat\Delta_n^N)\tr$ are the perturbation random vectors at instant $n$ with $\Delta_n^i,i=1,\ldots,N$, and $\widehat\Delta_n^i,i=1,\ldots,N$ being independent identically distributed (i.i.d), mean-zero random variables having finite inverse moments of order greater than 2 and satisfying the condition (A23) of section \ref{sec:2spsa-results}.

\section{Gradient estimate}
The SPSA estimate of the gradient $\nabla f(\theta_n)$ is given by
\begin{align}
\label{eq:grad-spsa}
\widehat\nabla_{(i)} f(\theta_n) =  \left[ \dfrac{y_n^+ - y_n^-}{2\delta_n \Delta_n^{(i)}}\right],
\end{align}
where the perturbations $\Delta_n^i$, $i=1,\ldots,N$ are i.i.d. and as mentioned above.

%Note that $E d_n^i = 0$, $E (d_n^i)^2 = 1+\epsilon$ and $E (d_n^i)^4 = \dfrac{(1+\epsilon)(1+(1+\epsilon)^3)}{(2+\epsilon)}$.

\section{Hessian estimate}
The $(i,j)$th entry of the Hessian estimate in this case is given by
\begin{align}
\label{eq:2spsa-ber}
&\left(\widehat H_n\right)_{ij} =  \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right).
\end{align}
%%%%%% Feedback
\section{Feedback term $\widehat \Psi_n$}
The $(i,j)$th term of the Hessian estimate $\widehat H_n$ can be simplified as follows: 
\begin{align}
\left(\widehat H_n \right)_{ij} & =  \left(\dfrac{y_n^+ + y_n^- - 2 y_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right) \nonumber \\
%& =  \left[\left(\dfrac{f(\theta_n+\delta_n \Delta_n+\delta_n \widehat \Delta_n) + f(\theta_n -\delta_n\Delta_n-\delta_n\widehat\Delta_n) - 2 f(\theta_n)}{\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}}\right) \right.\nonumber \\&\hspace{10em}+\left. \left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{\delta_n^2}\right)\right] \nonumber \\
& = \left( \frac{(\Delta_n+ \widehat \Delta_n)\tr \nabla^2 f(\theta_n) (\Delta_n+ \widehat \Delta_n)}{2\Delta_n^{(i)} \widehat \Delta_n^{(j)}} +  O(\delta_n^2) + \right.\left.\left(\dfrac{\xi_n^+ + \xi_n^- - 2 \xi_n}{2\delta_n^2 \Delta_n^{(i)} \widehat \Delta_n^{(j)}} \right)\right).\label{eq:h3}
%& = M_n \left(d_n\tr \nabla^2 f(\theta_n) d_n +  O(\delta_n^2) + O(\delta_n^{-2})\right). \label{eq:h0}
\end{align}
For the first term on the RHS above, note that
\begin{align}
&\E\left[\left.  \frac{(\Delta_n+ \widehat \Delta_n)\tr \nabla^2 f(\theta_n) (\Delta_n+ \widehat \Delta_n)}{2\Delta_n^{(i)} \widehat \Delta_n^{(j)}}  \right | \F_n\right] =\E\left[\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \Delta_n^{(m)}}{2\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \right.\nonumber \\&\hspace{2.5cm}+\left. \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)}}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} +\sum\limits_{l=1}^N\sum\limits_{m=1}^N \left. \frac{\widehat\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n)\widehat \Delta_n^{(m)}}{2\Delta_n^{(i)} \widehat\Delta_n^{(j)}}\right| \F_n\right].\label{eq:2spsaexp}
\end{align}
In analysing the RHS of the above expression, the first and the third terms are mean-zero (See Proposition 4.2 in \cite{bhatnagar2015simultaneous}):
\begin{align}
\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \Delta_n^{(m)}}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \right | \F_n\right] = 0 \hspace{0.2cm} and \hspace{0.2cm}\E\left[\left.\sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\widehat\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)}}{\Delta_n^{(i)} \widehat\Delta_n^{(j)} }\right | \F_n\right] = 0.
\end{align}
%\begin{align}
%\E\left[\left. \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^l \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^m}{\Delta_n^i \widehat\Delta_n^j} \right | \F_n\right] = \nabla^2_{ij} f(\theta_n) \hspace{0.3cm}a.s..
%\end{align}
From the above equations the  terms that are in expectation, denoted by  $\Psi_n^1(\nabla^2 f(\theta_n))$, can be written in matrix form as follows:
\begin{align}
\Psi_n^1(\nabla^2 f(\theta_n)) =\frac{1}{2} M_n \left[\Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + \widehat\Delta_n\tr \nabla^2 f(\theta_n) \widehat\Delta_n\right],
\end{align}
where $M_n = [1/\Delta_n^{(1)}, \ldots ,1/\Delta_n^{(N)}]\tr [1/\Delta_n^{(1)}, \ldots ,1/\Delta_n^{(N)}]$.
Now consider the second term in the RHS of \eqref{eq:2spsaexp}. It can be re-written as follows:
\begin{align}
\E\left[\left. \sum\limits_{l=1}^N\sum\limits_{m=1}^N \frac{\Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)}}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \right | \F_n\right] = \E\left[\left.\nabla^2_{ij} f(\theta_n) + \frac{1}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)} \right | \F_n\right]
\end{align}
In analysing the RHS of the above expression, the second term in the expectation  is mean-zero term:  
\begin{align}
\E\left[\left. \frac{1}{\Delta_n^{(i)} \widehat\Delta_n^{(j)}} \mathop{\sum_{l=1}^N \sum_{m=1}^{N}}_{lm \neq ij} \Delta_n^{(l)} \nabla^2_{lm} f(\theta_n) \widehat\Delta_n^{(m)} \right | \F_n\right] = 0.
\end{align}
The term on the LHS above, can be denoted by $\Psi_n^2(\nabla^2 f(\theta_n))$, and can be written in  matrix form as follows:
\begin{align}
\Psi_n^2(\nabla^2 f(\theta_n)) = \widehat N_n\tr \nabla^2 f(\theta_n) N_n  + \widehat N_n\tr \nabla^2 f(\theta_n) + \nabla^2 f(\theta_n) N_n.
\end{align}
where $N_n,\widehat N_n$ are defined as follows: $N_n = \Delta_n \left[\frac{1}{\Delta_n^{(1)}},\ldots,\frac{1}{\Delta_n^{(N)}}\right] - I_N$ and $\widehat N_n = \widehat \Delta_n \left [\frac{1}{\widehat\Delta_n^{(1)}},\ldots,\frac{1}{\widehat \Delta_n^{(N)}}\right] - I_N$ and $I_N$ is  identity matrix of size $N \times N$.
From the foregoing, the per-iteration Hessian estimate $\widehat H_n$ can be re-written as follows:
\begin{align}
\widehat H_n = & \nabla^2 f(\theta_n) + \Psi_{n}(\nabla^2 f(\theta_n)) +  O(\delta_n^2) + O(\delta_n^{-2})  \label{eq:spsahnhat}
\end{align}
where, for any matrix $H$, 
\begin{align}
\Psi_{n}(H) &=  \Psi_{n}^{1}(H) + \Psi_{n}^{2}(H)\nonumber\\
&=  \frac{1}{2}M_n \left[\Delta_n\tr H \Delta_n + \widehat\Delta_n\tr H \widehat\Delta_n\right] + \widehat N_n\tr H N_n + \widehat N_n\tr H + H N_n.\label{eq:spsapsi}
\end{align}
Given that we operate in a simulation optimization setting, which implies $\nabla^2 f$ is not known, we construct the feedback term $\widehat \Psi_n$ in \eqref{eq:2spsa-3-ih} by using $\overline H_{n-1}$ as a proxy for $\nabla^2 f$, i.e.,
\begin{align}
\widehat \Psi_n = \Psi_{n} (\overline H_{n-1}).
\label{eq:psinhatspsa3}
\end{align}

\paragraph{\textbf{Optimizing the step-sizes }$b_n$}

The optimal choice for $b_n$ in \eqref{eq:2spsa-3-ih}  is the following:
\begin{align}
\label{eq:spsawieghts}
b_i  = \delta_i^{4}/\sum\limits_{j=0}^{i} \delta_j^{4}.
\end{align}
The main idea behind the above choice is provided below.
From \eqref{eq:spsahnhat}, we can infer that
\begin{align*}
\E \parallel \widehat H_n \parallel^2 \le \dfrac{C}{\delta_n^4} \text{ for some } C<\infty. 
\end{align*} 
This is because the third term in \eqref{eq:spsahnhat} vanishes asymptotically, while the fourth term there dominates asymptotically. Moreover, the noise factors in the fourth term in \eqref{eq:spsahnhat} are bounded above due to (A28)  and independent of $n$, leaving the $\delta_n^2$ term in the denominator there.

So, the optimization problem to be solved at instant $n$ is as follows:
\begin{align}
\sum \limits_{i=0}^{n} (\tilde b_i)^2 \delta_i^{-4}, \text{ subject to} \label{eq:spsawn-opt}\\
\tilde b_i \geq 0 ~\forall i \text{ and }\sum \limits_{i=0}^{n} \tilde b_i = 1.
\end{align}
The optimization variable $\tilde b_i$ from the above is related to the Hessian recursion \eqref{eq:2spsa-3-ih} as follows:
\begin{align}
\label{eq:spsahess}
\overline H_n = \sum\limits_{i=0}^{n} \tilde b_i(\widehat H_i -\widehat \Psi_i).
\end{align}
The solution to \eqref{eq:spsawn-opt} is achieved for $\tilde b_i^* = \delta_i^{4}/\sum \limits_{j=0}^{n} \delta_j^{4}, i=1,\ldots,n$. The optimal choice $\tilde b_i^*$ can be translated to the step-sizes $b_i$, leading to \eqref{eq:spsawieghts}.

\section{Convergence analysis for 2SPSA-3-IH}
\label{sec:2spsa-results}
We make the same assumptions as those used in the analysis of \cite{prashanth2015rdsa}, with a few minor alterations. The assumptions are listed below:
\begin{pvn}
\item  The function
$f$ is four-times differentiable\footnote{Here $\nabla^4 f(\theta) = \dfrac{\partial^4 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the fourth derivate of $f$ at $\theta$ and $\nabla^4_{i_1, i_2, i_3, i_4} f(\theta)$ denotes the $(i_1, i_2, i_3, i_4)$th entry of $\nabla^4 f(\theta)$, for $i_1, i_2, i_3,i_4=1,\ldots, N$.} with $\left|\nabla^4_{i_1 i_2 i_3 i_4} f(\theta) \right| < \infty$, for $i_1, i_2, i_3,i_4=1,\ldots, N$ and for all $\theta\in \R^N$.

%\item  For some $\rho>0$  and almost all $\theta_n$, the function $f$ is four-times differentiable with a uniformly (in $n$) bounded fourth derivative for all $x$ such that $\left\| \theta_n - x\right\| \le \rho$.

\item For each $n$ and all $\theta$, there exists a $\rho>0$ not dependent on $n$ and $\theta$, such that $(\theta-\theta^*)\tr \bar f_n(\theta) \ge \rho \left\| \theta_n - \theta\right\|$, where $\bar f_n(\theta) = \Upsilon(\overline H_n)^{-1} \nabla f(\theta)$.

\item $\{\xi_n, \xi_n^+,\xi_n^-, n=1,2,\ldots\}$ are such that, for all $n$, $\E\left[\left. \xi_n^+ + \xi_n^- - 2 \xi_n \right| \F_n, \Delta_n, \widehat \Delta_n \right] = 0$, where $\mathcal{F}_n = \sigma(\theta_m,m < n)$ denotes the underlying sigma-field.

\item $\{\Delta_n^i,\widehat\Delta_n^i, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d independent of $\F_n$ and for some $\alpha_3 > 0$ and for all $n,l$ $|\Delta_{nl}|\le \alpha_3$, $\Delta_{nl}$ is symmetrically distributed about 0, $\Delta_{nl}$ are mutually independent across $n$ and $l$ and they satisfy $\E (\Delta_{nl}^{-2})$, $\E (\widehat\Delta_{nl}^{-2}) \le \alpha_3$.

\item  The step-sizes $a_n$ and perturbation constants $\delta_n$ are positive, for all $n$ and satisfy
$$\hspace{-1.7em} a_n, \delta_n \rightarrow 0\text{ as } n \rightarrow \infty, 
\sum_n a_n=\infty \text{ and } \sum_n \left(\frac{a_n}{\delta_n}\right)^2 <\infty.$$

\item For each $i=1,\ldots,N$ and any $\rho>0$, 
$P(\{ \bar f_{ni} (\theta_n) \ge 0 \text{ i.o}\} \cap \{ \bar f_{ni} (\theta_n) < 0 \text{ i.o}\} \mid \{ |\theta_{ni} - \theta^*_i| \ge \rho\quad \forall n\}) =0.$

\item The operator $\Upsilon$ satisfies $\delta_n^2 \Upsilon(H_n)^{-1} \rightarrow 0$ a.s. and  $E(\left\| \Upsilon(H_n)^{-1}\right\|^{2+\zeta}) \le \rho$ for some $\zeta, \rho>0$.

\item For any $\tau >0$ and nonempty $S \subseteq \{1,\ldots,N\}$, there exists a $\rho'(\tau,S)>\tau$ such that 
$$ \limsup_{n\rightarrow \infty} \left| \dfrac{\sum_{i \notin S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}{\sum_{i \in S} (\theta-\theta^*)_i \bar f_{ni}(\theta)}               \right| < 1 \text{ a.s.}$$
for all $|(\theta-\theta^*)_i| < \tau$ when $i \notin S$ and $|(\theta-\theta^*)_i| \ge \rho'(\tau,S)$ when $i\in S$.
\item For some $\alpha_7, \alpha_8, \alpha_9 >0$ and for all $n,l,m$, $\E {\xi_n}^{2} \le \alpha_7$, $\E {\xi_n^{\pm}}^{2} \le \alpha_7$, $\E f(\theta_n)^{2} \le \alpha_8$,  $\E f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2},\E f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2} \le \alpha_8$, 
%$\E \left[ f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[| f(\theta_n + \delta_n \Delta_n + \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_9} \mid \F_n\right]$,
%$\E \left[ f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)^{2}/ (\Delta_{nl}\widehat\Delta_{nm})^2 \mid \F_n\right]$,
$\E \left[ |f(\theta_n - \delta_n \Delta_n - \delta_n \widehat \Delta_n)/ (\Delta_{nl}\widehat\Delta_{nm})|^{2+\alpha_9} \mid \F_n\right]$,
 $\E\left[\left. (\xi_n^+ + \xi_n^- - 2 \xi_n)^2/(\Delta_{nl}\widehat\Delta_{nm})^2 \right| \F_n \right] \le \alpha_8$ 
and $\E \left(\left\| \Upsilon(\overline H_n) \right\|^2 \mid \F_n\right) \le \alpha_8$. 
\item  $\delta_n = \frac{\delta_0}{(n+1)^{\varsigma}}$, where $\delta_0 > 0$ and $0 < \varsigma \le 1/8$.
\end{pvn}
The reader is referred to Section II-B of \cite{prashanth2015rdsa} for a detailed discussion of the above assumptions. We remark here that (A20)-(A27) are identical to those in \cite{prashanth2015rdsa}, while (A28) and (A29) introduce minor additional requirements on $\left\| \Upsilon(\overline H_n) \right\|^2$ and $\delta_n$, respectively and these are inspired from \cite{spall-jacobian}.

\begin{lemma}(\textbf{2SPSA-3-IH Bias in Hessian estimate})
\label{lemma:2spsa-bias}
Under (A20)-(A29), with $\widehat H_n$ defined according to  \eqref{eq:2spsa-ber} , we have a.s. that\footnote{Here $\widehat H_n(i,j)$ and $\nabla^2_{i,j}f(\cdot)$ denote the $(i,j)$th entry in the Hessian estimate $\widehat H_n$ and the true Hessian $\nabla^2 f(\cdot)$, respectively.}, for $i,j = 1,\ldots,N$,
\begin{align}
\left|\E\left[
\left. \widehat H_n(i,j) \right| \F_n \right] - \nabla^2_{ij} f(\theta_n)\right| = O(\delta_n^2).
\end{align} 
\end{lemma}
\begin{proof}
See Proposition 4.2 in \cite{bhatnagar2015simultaneous}.
\end{proof}

\begin{theorem}(\textbf{2SPSA-3-IH Strong Convergence of Hessian})
\label{thm:2spsa-H}
Under (A20)-(A29), we have that 
$$\theta_n \rightarrow \theta^*, \overline H_n \rightarrow \nabla^2 f(\theta^*) \text{ a.s. as } n\rightarrow \infty.$$ 
In the above, $\theta_n$ and $\overline H_n$ are updated according to \eqref{eq:2spsa-3-theta} and \eqref{eq:2spsa-3-ih}, respectively, $\widehat H_n$ defined according to \eqref{eq:2spsa-ber} and the step-sizes $b_n$ are chosen as suggested in \eqref{eq:spsawieghts}. 
\end{theorem}
\begin{proof}
The proof is similar to the proof of the Theorem \ref{thm:2rdsa-H} and is therefore omitted.
\end{proof}

We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
\begin{pvn}
\item  $f$ is quadratic and $\nabla^2 f(\theta^*) > 0$. 
\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and \\$\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
\end{pvn}
\begin{theorem}(\textbf{2SPSA-3-IH Quadratic case - Convergence rate})
\label{thm:spsaquad-bound}
Assume (A23), (A29), (A30) and (A31) and also that the setting is noise-free. 
Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(\theta^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
\begin{align}
\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
\label{eq:spsaquad-bigo}
\end{align}
\end{theorem}
\begin{proof}
Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:spsahnhat} as follows:
\begin{align}
 \widehat H_n &=    \nabla^2 f(\theta_n) +\Psi_{n}(\nabla^2 f(\theta_n))\nonumber\\
 &=    \nabla^2 f(\theta^*) +\Psi_{n}(\nabla^2 f(\theta^*))\nonumber\\
 &=    H^* +\Psi_{n}(H^* ),\label{eq:spsahnhat-ext}
\end{align}
where $\Psi_n(H)$ is defined in \eqref{eq:psinhatspsa3}.
The proof involves the following steps:
\begin{description}
  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
  \item[Step 2:] We unroll the recursion \eqref{eq:2spsa-3-ih} and then derive a convenient representation for the  $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
  \item[Step 3:] We derive the main result in \eqref{eq:spsaquad-bigo} using a proof by contradiction. 
\end{description}
\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
This part of the proof follows along similar lines as  part (i) in Theorem 3 of \cite{spall-jacobian}.\\
By rewriting the \eqref{eq:2spsa-3-ih} in the following form
\begin{align}\label{eq:hhatexpanded}
\overline H_{n} &= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - \widehat H_n + \widehat \Psi_n)\nonumber\\
&= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - H^* + \widehat \Psi_n - \Psi_n (H^*)).
\end{align}
The above equation can be thought as a stochastic approximation analogue of \eqref{eq:2spsa-3-ih}. The above stochastic approximation algorithm is aimed at finding the roots of the equation $H-H^*=0$. In \eqref{eq:hhatexpanded}, $\E (\widehat \Psi_n - \Psi_n (H^*) | \F_n) = 0 \,\,\ a.s$.  Given a Lyapunov function, one can  get the conditions required to show  the mean square convergence of the \eqref{eq:hhatexpanded} from \cite{nevel1973stochastic}, pp.92-94. It is easy to see that $V(H) = \frac{1}{2} (H-H^*)\tr(H-H^*)$ indeed serves as Lyapunov function. From part (i) in Theorem 3 of \cite{spall-jacobian}, one can check that our algorithm indeed satisfies those conditions,  which in turn implies that $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0 \,\,\,a.s.$ as $n\rightarrow \infty$.

\noindent\textbf{Step 2: Representation of $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$} \\
 From \eqref{eq:2spsa-3-ih} and \eqref{eq:spsahnhat-ext},  we have
 \begin{align}
 \Lambda_n &= \Lambda_{n-1} - b_n ( \overline H_{n-1} - \hat H_n + \hat \Psi_n) \nonumber\\
&= (1-b_n) \Lambda_{n-1} - b_n (H^* + \hat \Psi_n - \hat H_n ) \nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n (\hat \Psi_n - \Psi_n(H^*)) \nonumber\\&= (1-b_n) \Lambda_{n-1} - b_n \Psi_n(\Upsilon(\overline H_{n-1})) + b_n \Psi_n(H^*) \nonumber\\
&= (1-b_n) \Lambda_{n-1} - b_n\Psi_n(\Lambda_{n-1}'),\label{eq:step1last}
 \end{align}
where $\Lambda_n' = \Upsilon(\overline H_n) - H^*$.
The equality in \eqref{eq:step1last} follows from \eqref{eq:psi}. 
Unrolling the recursion \eqref{eq:step1last}, we obtain 
 \begin{align}\label{lambda-exp}
  \Lambda_n  = \left[ \prod_{k=1}^n (1-b_k) \right]\Lambda_0 
&- \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right] b_k \Psi_k(\Lambda_{k-1}')  \hspace{0.2cm} a.s.
 \end{align}
 %%(note : $\prod_{j=n+1}^n (1-b_j) = 1$ for all n).
 %%Let us  characterise $trace[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{lambda-exp}. From independence of $d_k$ along k, \eqref{lambda-exp} represent martingale difference sequence, leading to 
It follows from \eqref{lambda-exp} that 
 \begin{align}
  \E (\Lambda_n \tr \Lambda_n) =  &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \E (\Lambda_0 \tr \Lambda_0) + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \E (\Psi_k(\Lambda_{k-1}') \tr \Psi_k(\Lambda_{k-1}')).\label{eq:exp-lambda2}
 \end{align}
The equality above uses the fact that $\E(\Psi_k(\Lambda_{k-1}'))=0$, which gets rid of the corresponding cross terms with the first term on RHS of \eqref{lambda-exp}. % %
We now characterize $\text{trace}[\E(\Lambda_n\tr \Lambda_n)]$ using \eqref{eq:exp-lambda2} as follows:
%From the independence of $d_k$'s and $\Lambda_{k-1}'$ and using Cauchy-Schwartz inequality we get 
 \begin{align}\label{eq:trace}
 \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &\left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right).
 \end{align}
 % \begin{align}\label{eq:trace}
% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq \left[ \prod_{k=1}^n (1-b_k) \right]^2 \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] + \nonumber\\ &+ \sum_{k=1}^n \left[\prod_{j=k+1}^n (1-b_j)\right]^2 b_k^2   \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
% \end{align}
whereas in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a  fashion and then returns the trace of the resulting $N\times N$ matrix.
As in the proof of Theorem 3 of \cite{spall-jacobian}, observe that $1-b_k = e^{-b_k}(1-O(b_k^2))$ and since $0 < b_k <1$, we have that the $O(b_k^2)$ term is strictly positive. 
Letting $\Gamma_{ij} = \sum_{k=i}^j b_k$ with $\Gamma_{nn} = 1$ and $\beta_{kn} = \left[\prod_{i=k+1}^n (1- O(b_i^2))\right]^2$, we can simplify \eqref{eq:trace} as follows:
\begin{align}\label{eq:wij}
 \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] = &e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  + e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right).
 \end{align} 
% \begin{align}\label{eq:wij}
% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] \leq e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \,\,\tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \nonumber\\&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ e^{-2 \Gamma_{1n}} \sum_{k=1}^n e^{2 \Gamma_{1k}} \beta_{kn} b_k^2  \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2} \nonumber\\&  \times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
% \end{align} 
Comparing the sum with integrals, we obtain
\begin{align*}
\Gamma_{ij} = \int_i^j \frac{b_0}{x^r} dx + O(1) &= \left(\frac{b_0}{1-r}\right)(j^{1-r}-i^{1-r}) +O(1),
\end{align*}
where we have used the facts that $0 < b_k < 1, \forall k \ge 2$ and 
$\sum_{k=i}^j b_k \to \infty$ as $j-i \to \infty$ since 
$b_k = b_0/k^r$ with $r > 0.5$.
Observing that $\beta_{kn}$ are uniformly upper-bounded, say by $\bar \beta_n$, we have
\begin{align}\label{eq:thmwts}
 & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
&\hspace{3cm}+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right).
\end{align} 
%\begin{align}\label{eq:thmcomb}
%& \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \text{trace}\left[\E [((\Phi_k(H^*) - H^*) - 2 \Psi_k(\Lambda_{k-1}))\tr  (\Phi_k(H^*) - H^*)]\right] .
%\end{align} 
%
%%\begin{align}\label{eq:thmwts}
%% & \text{trace} \left[\E (\Lambda_n \tr \Lambda_n)\right] =  e^{-2 \Gamma_{1n}} \beta_{0n} \text{trace} \left[\E (\Lambda_0 \tr \Lambda_0)\right]  \nonumber\\ 
%%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}} \nonumber\\
%%& \hspace{3.5cm} \times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)  \nonumber\\ 
%%&+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\
%%&  \hspace{0.5cm}\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right] \nonumber\\ &+ \bar \beta_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \times \frac{b_0^2}{k^{2 r}}  \nonumber\\&\hspace{3.5cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2}\nonumber\\&\times \text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]^{1/2}.
%%\end{align}
%\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
%In comparison to Eq. (7.6) in \cite{spall-jacobian} that corresponds to \eqref{eq:thmwts} for the N-SPSA-4 setting, there are two extra terms - the third and the fourth - in the RHS of \eqref{eq:thmwts}. 
%Consider the third term on the RHS of \eqref{eq:thmwts}. As a consequence of (C9) and the construction of random perturbations $d_k$, it can be seen that $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$ is uniformly bounded above, say by $\top$, independent of $k$. Thus, the third term is equivalent to the following:
%\begin{align}
% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \top.
%\label{eq:term3}
%\end{align}
%Now by moving out side exponential term to inside and changing the summation to integration, we can rewrite the above equation as follows
%\begin{align}
%  \overline{\overline{c_n}}  \int_1^n  \frac{1}{k^{2 r}} .
%\label{eq:term3}
%\end{align}
%Where $\overline{\overline{c_n}}$ is a constant. By above equation we can conclude that the third term is of $O(\frac{1}{k^{2 r -1}})$
%%Now, observe that $\lim_{n\rightarrow \infty} \sum_{k=1}^n \frac{1}{k^{2 r}} < \infty$ as $1/2 < r < 1$ and $e^{2 b_0 k^{1-r}/(1-r)}$ diverges as $n\rightarrow \infty$ and hence, by Kronecker's Lemma, the term in \eqref{eq:term3} vanishes asymptotically. 
%%
%%Now consider the fourth term. By using the uniform upper-bound $\top$ on $\text{trace}\left[\E ((\Phi_k(H^*) - H^*) \tr  (\Phi_k(H^*) - H^*))\right]$, we can rewrite  the fourth term in \eqref{eq:thmwts} as follows: 
%% \begin{align}\label{eq:fourthterm}
%% \bar c_n e^{-2 b_0 n^{1-r}/(1-r)} \sum_{k=1}^n e^{2 b_0 k^{1-r}/(1-r)}  \frac{b_0^2}{k^{2 r}} \sqrt{\top} \nonumber \\ \hspace{3cm}\times \tau\left(\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')\right)^{1/2},
%% \end{align}
%%where, as in the proof of Theorem 3 of \cite{spall-jacobian}, $\tau(\cdot)$ transforms the $\E (\Lambda_{k-1}' \otimes \Lambda_{k-1}')$ in a  fashion and then returns the trace of the resulting $N\times N$ matrix.

\noindent\textbf{Step 3: The big-O result on $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$ convergence} \\
We will use similar argument in part (iii) in Theorem 3 of \cite{spall-jacobian} to show the claim \eqref{eq:spsaquad-bigo}. The idea here is to show that the contradiction exists if RHS of \eqref{eq:spsaquad-bigo} exhibits  any slower rate of decay. Let us assume the following slower rate of decay for the LHS of \eqref{eq:spsaquad-bigo}.
\begin{align}
\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = d(n) e^{-2b_0n^{1-r} / {1-r}},
\label{eq:spsaquad-bigocont}
\end{align}
where $d(n) = o(e^{2b_0n^{1-r} / {1-r}})$. Now we will to show that there is a contradiction. By using the series of arguments used in part (iii) in Theorem 3 of \cite{spall-jacobian} and applying the condition (A31) we get the following: 
\begin{align}
\E (\Lambda_{n}' \otimes \Lambda_{n}') = \E (\Lambda_{n} \otimes \Lambda_{n}) + o(d(n) e^{-2b_0n^{1-r} / {1-r}}).
\label{eq:spsaquad-bigocont2}
\end{align}
Now from the above equation we can write 
\begin{align}
\E (\Lambda_{n-1}' \otimes \Lambda_{n-1}') = O (\text{trace}[\E (\Lambda_{n-1} \tr \Lambda_{n-1})]). 
\label{eq:spsaquad-bigocont3}
\end{align}

Now we can use the result in the RHS of equation \eqref{eq:thmwts} to show the contradiction. We consider two cases to show the contradiction. The first is where $\frac{d(n)}{n^{2r}} = O(1)$ and the other is $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$. We show contradiction for the both cases separately. Let us consider the first case, i.e., $\frac{d(n)}{n^{2r}} = O(1)$, Then
\begin{align}
\text{RHS of \eqref{eq:thmwts}} &= O(1) e^{-2 \Gamma_{1n}} +  O(1) e^{-2 b_0 n^{1-r}/(1-r)} \sum_{i=1}^n \frac{d(i)}{i^{p}} \frac{b_0^2}{i^{2r-p}}\nonumber \\
&= e^{-2b_0n^{1-r} / {1-r}} \left[ O(1) + O(1) \sum_{i =1}^n \frac{1}{i^{2r-p}}\right],
\label{eq:spsaquad-bigocont4}
\end{align}
where there exist  $0 < p \le 2r$, $0 < \epsilon < 2r -1$ such that $\frac{d(n)}{i^{p}} = O(1)$ and $lim sup_{n \to \infty} \frac{d(n)}{n^{p-\epsilon}} = \infty$. In \eqref{eq:spsaquad-bigocont4}, the first equality uses \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont3}.

Now there exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{p-\epsilon}} = \infty$ as $n \to \infty$. Now LHS of \eqref{eq:thmwts} satisfies $LHS(i_n) \ge C i_n^{p-\epsilon} e^{-2b_0 i_n^{1-r} / {1-r}}$ for some $C > 0$. Now we have $\frac{LHS(i_n)}{RHS(i_n)} \to 0$ because  $0 < \epsilon < 2r -1$. This leads to a contradiction since $LHS(n) = RHS(n)$ for all $n$.

Now consider the second case where $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$. There exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{2r}} \ge \frac{d(i)}{i^{2r}}$ as $i \le i_n$. Then for some $C >0$ the RHS of \eqref{eq:thmwts} satisfies the following:
\begin{align}
\text{RHS}(i_n)\,\,\,  \text{of}  \,\,\,\eqref{eq:thmwts} &= O(1) e^{-2 \Gamma_{1i_n}} +  O(1) e^{-2 b_0 i_n^{1-r}/(1-r)} \left[\sum_{i=1}^{i_n} e^{2 b_0 i^{1-r}/(1-r)} \frac{b_0^2}{i^{2r}} \E (\Lambda_{n} \otimes \Lambda_{n}) \right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + O(1) \sum_{i =1}^{i_n} \frac{d(i)}{i^{2r}}\right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + \frac{d(i_n)}{{i_n}^{2r}}\right],
\label{eq:spsaquad-bigocont5}
\end{align}
where $\text{RHS}(i_n)$ denotes, RHS after the algorithm is iterated for $i_n$ times. From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} after $i_n$ iterations is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$. By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS(i_n)}{LHS(i_n)}$, it turns out that $\frac{RHS(i_n)}{LHS(i_n)} \to 0$ as $n \to \infty$. This results in a contradiction since $LHS(n) = RHS(n)$ for all $n$. Hence the claim $\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}})$ follows.
\end{proof}