### 2017-05-01 response.tex
response to reviwer comments

%\documentclass[oneside,8pt,onecolumn,openright]{IIScthesisPSnPDF}
\documentclass[12pt]{article}

\usepackage{paralist}
\let\labelindent\relax

\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{colortbl}
%\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.12}
\usepackage{todonotes}
\usepackage{nicefrac}
\usetikzlibrary{positioning,arrows,calc,patterns,backgrounds}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{caption}
\usepackage{subcaption}

\newcommand{\tabl}[2]{\begin{tabular}{#1} #2 \end{tabular}}
\newcommand{\ith}[2][th]{$#2^{\text{#1}}$}

%\newcommand{\cP}{\mathcal{P}}
%\newcommand{\cS}{\mathcal{S}}
%\newcommand{\R}{\mathbb{R}}
%\newcommand{\E}{\mathbb{E}}
%\renewcommand{\S}{\mathcal{S}}
%\newcommand{\D}{\mathcal{D}}
%\newcommand{\A}{\mathcal{A}}
%\newcommand{\B}{\mathcal{B}}
%\newcommand{\X}{\mathcal{X}}
\newcommand{\F}{\mathcal{F}}
%\newcommand{\G}{\mathcal{G}}
\newcommand{\N}{\mathcal{N}}
%\newcommand{\Z}{\mathcal{Z}}
%\newcommand{\T}{\mathcal{T}}
%\renewcommand{\O}{\mathcal{O}}
%\renewcommand{\P}{\mathbb{P}}

\newcommand{\tr}{^\mathsf{\scriptscriptstyle T}}
\newcommand{\tpi}{\Pi}
\newcommand{\tgamma}{\tilde{\gamma}}
\newcommand{\ttheta}{\tilde{\theta}}
\newcommand{\tz}{\tilde{z}}

\renewcommand{\l}{\left\|}
\renewcommand{\r}{\right\|}
\newcommand{\vml}{\left\|}
\newcommand{\vmr}{\right\| _{\Psi}}
\newcommand{\ml}{\left\|}
\newcommand{\mr}{\right\|_2}

\newcommand{\e}{\exp}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{fancyhdr}
\usepackage[numbers]{natbib}
\usepackage{cleveref}
\usepackage{wrapfig}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{mathrsfs}
% \usepackage{sfgame}
\usepackage{color}
\usepackage[compact]{titlesec}
\usepackage{mdwlist}
\usepackage{varwidth}
\usepackage[vertfit]{breakurl}
\usepackage{datetime}
\usepackage{pdfpages}
% \usepackage{biblatex}
\usepackage{setspace}
\title{\textbf{Response to Reviewer Comments}}
\author{}
\date{}
\begin{document}
%\Huge
%\begin{center}
%Panel of Experts
%\end{center}
\maketitle
\begin{tabular}{ll}
Name of the Candidate & \textbf{Danda Sai Koti Reddy}\\ 
SR Number &\textbf{04-04-00-10-21-14-1-11609} \\
Title of the Thesis & \textbf{Stochastic Newton methods with   } \\
 & \textbf{ enhanced Hessian estimation.}\\
 Research Supervisor  & \textbf{Prof. Shalabh Bhatnagar} \\
Degree Registered  &  \textbf{M.Sc(engg)} \\
Department  & \textbf{Computer Science and Automation} \\
Institute  & \textbf{Indian Institute of Science, Bangalore }
\end{tabular}

\vspace{2cm}

\section{Reviewer Suggestions}
The questions are presented in \textit{italicized} font and the answers in normal font. The response to the reviewer questions are as under.\\

\textit{\textbf{Suggestion (i): } A key early work in this direction merits mention: \\  \\Anbar, Dan. ``A stochastic Newton-Raphson method." Journal of Statistical Planning and Inference 2.2 (1978): 153-163.\\ \\ A sleeker `ODE' based convergent argument might be facilitated by:\\ \\  Smale, Steve. ``A convergent process of price adjustment and global Newton methods." Journal of Mathematical Economics 3.2 (1976): 107-120.\\ \\}
 \textbf{Response:}\\
We have mentioned the references given in suggestion in the section 1.5, titled ``Hessian estimation". Which is given as below \\ \\
\textbf{1.5 Hessian estimation} \\
Stochastic Newton methods can counter the ill-conditioning problem of the objective $f$ as they incorporate second-order information into the update iteration given by   
\begin{align}
\label{eq:newton}
\theta_{n+1} = \theta_n - a_n (\overline H_n)^{-1}\widehat\nabla f(\theta_n), 
\end{align}
where $a_n$ is the step-size that satisfies standard stochastic approximation conditions (see (A12) in Section 2.6), and $\widehat\nabla f(\theta_n)$ and $\overline H_n$ are estimates of the gradient and Hessian, respectively. Thus, \eqref{eq:newton} can be considered as  the stochastic version of the well-known Newton method for optimization. Stochastic Newton methods are often more accurate than simple gradient search schemes. A key early work in stochastic Newton methods  is given in \cite{anbar1978stochastic} and a sleek `ODE' based convergence argument is provided in \cite{smale1976convergent}.\\ \\
\textit{\textbf{Suggestion (ii):} I would also appreciate any comments on the following issue: Even in the deterministic scenario, Newton scheme is not popular due to the expensive matrix inversion, and quasi-Newton schemes given by the Broyden family are used instead. Why not develope stochastic approximation versions of the latter?  }\\ \\
 \textbf{Response:}\\
 We have mentioned this as an interesting future direction in chapter 6, titled ``Conclusions and Future work". Which is given below \\
 3. Newton scheme is not popular due to expensive matrix inversion at each parameter update step even in deterministic scenario. Quasi-Newton schemes given by the Broyden family may be used instead. As a future work, it would be interesting to develop stochastic approximation versions of the quasi-Newton schemes.

\bibliographystyle{plain}
\bibliography{refThesis}

\end{document}